Researchers collaborate with same-gendered colleagues more often than expected across the life sciences

### 2019-03-04 manuscript/homophily_manuscript.tex

\begin{document}
\maketitle
\begin{abstract}
Evidence suggests that women in academia are hindered by conscious and
unconscious biases, and often feel excluded from formal and informal
opportunities for research collaboration. In addition to ensuring
fairness and helping to redress gender imbalance in the academic
workforce, increasing women's access to collaboration could help
scientific progress by drawing on more of the available human capital.
Here, we test whether researchers tend to collaborate with same-gendered
colleagues, using more stringent methods and a larger dataset than in
past work. Our results reaffirm that researchers co-publish with
colleagues of the same gender more often than expected by chance, and
show that this `gender homophily' is slightly stronger today than it was
10 years ago. Contrary to our expectations, we found no evidence that
homophily is driven mostly by senior academics, and no evidence that
homophily is stronger in fields where women are in the minority.
Interestingly, journals with a high impact factor for their discipline
tended to have comparatively low homophily, as predicted if mixed-gender
teams produce better research. We discuss some potential causes of
gender homophily in science. \vspace{5mm}
\par\noindent \textbf{Keywords:} Gender bias, Homophily, Research
collaboration, Text mining, Women in STEM.
\end{abstract}

\maketitle

\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
\maketitle\unmarkedfntext{\textsuperscript{§}Organismal and Evolutionary Biology Research Programme,
Faculty of Biological and Environmental Sciences, University of Helsinki, Finland.}

@@

However, we believe that earlier studies of gender homophily were
hindered by a largely unacknowledged statistical issue that we will
refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with
the conceptually similar Whalund effect in population genetics {[}59{]}.
The Wahlund effect makes it deceptively difficult to infer gender-based
preferences simply by counting the number of same- and mixed-gender
coauthorships. Essentially, whenever coauthorship data are sampled from
two or more discrete sets of literature, which vary in the author gender
ratio and which are largely not connected by collaboration, the number
of same-gendered coauthors will be inflated. This can give the
impression that authors preferentially publish with same-gendered
colleagues even if no gender preferences exist, or if the true
preference is for opposite-gendered colleagues (`gender heterophily').
For example, a sample of literature containing a mixture of
bioinformatics and cell biology papers will probably contain an excess
of mostly-male and mostly-female author lists, simply because
researchers usually collaborate within their own discipline, and because
the author gender ratio is more male-biased in bioinformatics than in
cell biology {[}5{]}.

@@

\section{Results}\label{results}

\subsection{Gender homophily by discipline, time period, and authorship
position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}

@@

\subsection{Variance in homophily between
disciplines}\label{variance-in-homophily-between-disciplines}

@@

\subsection{Relationship between gender homophily and number of
authors}\label{relationship-between-gender-homophily-and-number-of-authors}

Papers with two authors had significantly lower (but still positive)
\(\alpha'\) values relative to papers with more than two authors, on
average (\autoref{author_number}; statistical results in Online
Supplementary Material). Papers with 3, 4 or \(\ge5\) authors had
essentially identical average \(\alpha'\) values. The variance in
\(\alpha'\) across journals was also a little higher for 2-authors
papers compared to the remainder (\autoref{author_number}), though part
of this variance is due to the reduced sample size (in terms of number
of authors) for the 2-author papers. One possible explanation for this
finding is that 2-authors papers might be more likely to have an author
list that is evenly split between career stages (e.g.~a postgraduate
student and their supervisor), increasing the chance that the authors
are mixed gender (see Figure 6). The result also suggests that the
processes responsible for gender homophily are similar in small (e.g.~3
author) and larger (5+ authors) collaborations (and across disciplines
where small versus large collaborations are the norm).

@@

\subsection{Relationship between gender homophily and gender
ratio}\label{relationship-between-gender-homophily-and-gender-ratio}

@@

\subsection{Relationship between journal impact factor and gender
homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}

@@

\subsection{Analysis accounting for differences in author gender ratio
between
countries}\label{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}

@@

\subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
the gender ratio differs between career
stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}

@@

\section{Discussion}\label{discussion}

@@

\section{Methods}\label{methods}

\subsection{The dataset}\label{the-dataset}

@@

\subsection{\texorpdfstring{Calculating \(\alpha\), the coefficient of
homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}

@@

Our test assumes that the expected value of \(\alpha\) is zero if
authors randomly assort, but for small datasets this assumption is not
always true. Essentially, this issue arises because a person cannot be
their own co-author. In a small dataset comprising \(m\) men and \(f\)
women, a man can co-author with \(m - 1\) men while women can co-author
with \(m\) men. Thus, the null expectation for \(alpha\) is a negative
number -- potentially a large one if \(m\) and \(f\) are very small.

@@

\subsection{Minimising the Wahlund effect: research discipline and time
period}\label{minimising-the-wahlund-effect-research-discipline-and-time-period}

@@

\subsection{Minimising the Wahlund effect: author country of
affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}

@@

\subsection{Calculating standardised journal impact
factor}\label{calculating-standardised-journal-impact-factor}

@@

\subsection{Statistical analysis}\label{statistical-analysis}

@@

\subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
the gender ratio differs between career
stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}

@@

\subsection{Data availability and
reproducibility}\label{data-availability-and-reproducibility}

@@

\section{Acknowledgements}\label{acknowledgements}

@@

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-Shaw_2012}{}
1. Shaw AK, Stanton DE. Leaks in the pipeline: separating demographic
inertia from ongoing gender differences in academia. Proceedings of the
Royal Society of London B. 2012;272: 3736--3741.

\hypertarget{ref-Lariviere_2013}{}
2. Larivière V, Ni C, Gingras Y, Cronin B, Sugimoto CR. Bibliometrics:
global gender disparities in science. Nature. 2013;504: 211--213.

\hypertarget{ref-West_2013}{}
3. West JD, Jacquet J, King MM, Correll SJ, Bergstrom CT. The role of
gender in scholarly authorship. PLoS ONE. 2013;8: e66212.

\hypertarget{ref-Elsevier_report}{}
4. Elsevier Report. Gender in the global research landscape.
elseviercom/research-intelligence/resource-library/gender-report. 2017;

\hypertarget{ref-Holman_2018}{}
5. Holman L, Stuart Fox D, Hauser CE. The gender gap in science: How
long until women are equally represented? PLoS Biology. 2018;16:
e2004956.

\hypertarget{ref-Wutte_2007}{}
6. Wutte M. Closing the gender gap. Nature. 2007;448: NJ101--NJ102.

\hypertarget{ref-Reuben_2014}{}
7. Reuben E, Sapienza P, Zingales L. How stereotypes impair women's
careers in science. Proceedings of the National Academy of Sciences.
2014;111: 4403--4408.

\hypertarget{ref-Trower_2002}{}
8. Trower CA, Chait RP. Faculty diversity: Why women and minorities are
underrepresented in the professoriate, and fresh ideas to induce needed
reform. Harvard Magazine. 2002;104: 33--37.

\hypertarget{ref-Umbach_2007}{}
9. Umbach PD. Gender equity in the academic labor market: An analysis of
academic disciplines. Research in Higher Education. 2007;48: 169--192.

\hypertarget{ref-Hosek_2005}{}
10. Hosek S, Cox AG, Ghosh-Dastidar B, Kofner A, Ramphal N, Scott J, et
al. Gender differences in major federal external grant programs. RAND
Corporation. 2005;

\hypertarget{ref-pohlhaus_2011}{}
11. Pohlhaus JR, Jiang H, Wagner RM, Schaffer WT, Pinn VW. Sex
differences in application, success, and funding rates for NIH
extramural programs. Academic Medicine. 2011;86: 759.

\hypertarget{ref-Zuckerman_1987}{}
12. Zuckerman H. Persistence and change in the careers of men and women
scientists and engineers. National Academy Press. 1987; 127--156.

\hypertarget{ref-Rosenfeld_1991}{}
13. Rosenfeld RA. Outcome analysis of academic careers. Review prepared
for the Office of Scientific and Engineering Personnel, National
Research Council. 1991;

\hypertarget{ref-Long_1993}{}
14. Long JS, Paul DA, Robert M. Rank advancement in academic careers:
Sex differences and the effects of productivity. American Sociological
Review. 1993; 703--722.

\hypertarget{ref-Hopkins_2013}{}
15. Hopkins AL, Jawitz JW, McCarty C, Goldman A, Basu NB. Disparities in
publication patterns by gender, race and ethnicity based on a survey of
a random sample of authors. Scientometrics. 2013;96: 515--534.

\hypertarget{ref-ODorchai_2009}{}
16. O'Dorchai S, Meulders D, Crippa F, Margherita A. She figures
2009--Statistics and indicators on gender equality in science.
Publications Office of the European Union. 2009;

\hypertarget{ref-Feldt_1986}{}
17. Feldt B. The faculty cohort study: School of medicine. Ann Arbor,
MI: Office of Affirmative Action. 1986;

\hypertarget{ref-Stack_2004}{}
18. Stack S. Gender, children and research productivity. Scientometrics.
2004;45: 891--920.

\hypertarget{ref-Lariviere_2011}{}
19. Larivière V, Vignola-Gagné E, Villeneuve C, Gélinas P, Gingras Y.
Sex differences in research funding, productivity and impact: an
analysis of Québec university professors. Scientometrics. 2011;87:
483--498.

\hypertarget{ref-Moss_2012}{}
20. Moss-Racusin CA, Dovidio JF, Brescoll VL, Graham MJ, Handelsman J.
Science faculty's subtle gender biases favor male students. Proceedings
of the National Academy of Sciences. 2012;109: 16474--16479.

\hypertarget{ref-Knobloch_2013}{}
21. Knobloch-Westerwick S, Glynn CJ, Huge M. Science faculty's subtle
gender biases favor male students. Science Communication. 2013;35:
603--625.

\hypertarget{ref-Lee_2005}{}
22. Lee S, Bozeman B. The impact of research collaboration on scientific
productivity. Social Studies of Science. 2005;35: 673--702.

\hypertarget{ref-Wuchty_2007}{}
23. Wuchty S, Jones BF, Uzzi B. The increasing dominance of teams in
production of knowledge. Science. 2007;316: 1036--1039.

\hypertarget{ref-Abramo_2009}{}
24. Abramo G, D'Angelo CA, Di Costa F. Research collaboration and
productivity: is there correlation? Higher Education. 2009;57: 155--171.

\hypertarget{ref-Lariviere_2015}{}
25. Larivière V, Gingras Y, Sugimoto CR, Tsou A. Team size matters:
Collaboration and scientific impact since 1900. Journal of the
Association for Information Science and Technology. 2015;66: 1323--1332.

\hypertarget{ref-Long_1992}{}
26. Long JS. Measures of sex differences in scientific productivity.
Social Forces. 1992;71: 159--178.

\hypertarget{ref-Bozeman_2011}{}
27. Bozeman B, Monica G. How do men and women differ in research
collaborations? An analysis of the collaborative motives and strategies
of academic researchers. Research Policy. 2011;40: 1393--1402.

\hypertarget{ref-Abramo_2013}{}
28. Abramo G, D'Angelo CA, Di Costa F. Gender differences in research
collaboration. Journal of Informetrics. 2013;7: 811--822.

\hypertarget{ref-Badar_2013}{}
29. Badar K, Hite JM, Badir YF. Examining the relationship of
co-authorship network centrality and gender on academic research
performance: The case of chemistry researchers in pakistan.
Scientometrics. 2013;94: 755--775.

\hypertarget{ref-Lewison_2001}{}
30. Lewison G. The quantity and quality of female researchers: A
bibliometric study of Iceland. Scientometrics. 2001;52: 29--43.

\hypertarget{ref-Webster_2001}{}
31. Webster BM. Polish women in science: A bibliometric analysis of
Polish science and its publications. Research Evaluation. 2001;10:
185--194.

\hypertarget{ref-Bozeman_2004}{}
32. Bozeman B, Corley E. Scientists' collaboration strategies:
implications for scientific and technical human capital. Research
Policy. 2004;33: 599--616.

\hypertarget{ref-Long_1990}{}
33. Long JS. The origins of sex differences in science. Social Forces.
1990;68: 1297--1316.

\hypertarget{ref-Fuchs_2001}{}
34. Fuchs S, Von Stebut J, Allmendinger J. Gender, science, and
scientific organizations in Germany. Minerva. 2001;39: 175--201.

\hypertarget{ref-Reskin_1978}{}
35. Reskin BF. Scientific productivity, sex, and location in the
institution of science. American Journal of Sociology. 1978;83:
1235--1243.

\hypertarget{ref-Wright_2003}{}
36. Wright AL, Schwindt LA, Bassford TL, Reyna VF, Shisslak PAS
Catherine M amd Germain, Reed KL. Gender differences in academic
advancement: Patterns, causes, and potential solutions in one U.S.
college of medicine. Social Forces. 2003;68: 1297--1316.

\hypertarget{ref-bleidorn2016age}{}
37. Bleidorn W, Arslan RC, Denissen JJ, Rentfrow PJ, Gebauer JE, Potter
J, et al. Age and gender differences in self-esteem -- a cross-cultural
window. Journal of Personality and Social Psychology. 2016;111: 396.

\hypertarget{ref-jagsi_2016}{}
38. Jagsi R, Griffith KA, Jones R, Perumalswami CR, Ubel P, Stewart A.
Sexual harassment and discrimination experiences of academic medical
faculty. JAMA. 2016;315: 2120--2121.

\hypertarget{ref-Martin_2014}{}
39. Martin JL. Ten simple rules to achieve conference speaker gender
balance. PLoS computational biology. 2014;10: e1003903.

\hypertarget{ref-Tower_2007}{}
40. Tower G, Julie P, Brenda R. A multidisciplinary study of
gender-based research productivity in the world's best journals. Journal
of Diversity Management. 2007;2: 23--32.

\hypertarget{ref-Jordan_2008}{}
41. Jordan CE, Clark SJ, Vann CE. Do gender differences exist in the
publication productivity of accounting faculty?. Journal of Applied
Business Research. 2008;24: 77--85.

\hypertarget{ref-Britton_2000}{}
42. Britton DM. The epistemology of the gendered organization. Gender
and Society. 2000;14: 418--434.

\hypertarget{ref-Reagans_2001}{}
43. Reagans R, Zuckerman EW. Networks, diversity, and productivity: The
social capital of corporate R\&D teams. Organization Science. 2001;12:
502--517.

\hypertarget{ref-Hong_2004}{}
44. Hong L, Page SE. Groups of diverse problem solvers can outperform
groups of high-ability problem solvers. Proceedings of the National
Academy of Sciences. 2004;101: 16385--16389.

\hypertarget{ref-Whittington_2008}{}
45. Whittington KB, Smith-Doerr L. Women inventors in context:
Disparities in patenting across academia and industry. Gender \&
Society. 2008;22: 194--218.

\hypertarget{ref-Bear_2011}{}
46. Bear JB, Woolley AW. The role of gender in team collaboration and
performance. Interdisciplinary Science Reviews. 2011;36: 46--153.

\hypertarget{ref-Herrera_2012}{}
47. Herrera R, Duncan PA, Green MT, Skaggs SL. The effect of gender on
leadership and culture. Global Business and Organizational Excellence.
2012;31: 37--48.

\hypertarget{ref-Campbell_2013}{}
48. Campbell LG, Mehtani S, Dozier ME, Rinehart J. Gender-heterogeneous
working groups produce higher quality science. PloS ONE. 2013; e79147.

\hypertarget{ref-Ferber_1980}{}
49. Ferber MA, Teiman M. Are women economists at a disadvantage in
publishing journal articles? Eastern Economic Journal. 1980;6:
1189--193.

\hypertarget{ref-McDowell_1992}{}
50. McDowell JM, Smith JK. The effect of gender-sorting on propensity to
coauthor: Implications for academic promotion. Economic Inquiry.
1992;30: 68--82.

\hypertarget{ref-Ghiasi_2015}{}
51. Ghiasi G, Larivière V, Sugimoto CR. On the compliance of women
engineers with a gendered scientific system. PloS ONE. 2015;10:
e0145931.

\hypertarget{ref-Crow_2015}{}
52. Crow MS, Smykla JO. An examination of author characteristics in
national and regional criminology and criminal justice journals,
2008-2010: Are female scholars changing the nature of publishing in
criminology and criminal justice? American Journal of Criminal Justice.
2015;40: 441--455.

\hypertarget{ref-Fahmy_2017}{}
53. Fahmy C, Young JT. Gender inequality and knowledge production in
criminology and criminal justice. Journal of Criminal Justice Education.
2017;28: 285--305.

>>>>

\begin{document}
\maketitle
\begin{abstract}
Evidence suggests that women in academia are hindered by conscious and
unconscious biases, and often feel excluded from formal and informal
opportunities for research collaboration. In addition to ensuring
fairness and helping to redress gender imbalance in the academic
workforce, increasing women's access to collaboration could help
scientific progress by drawing on more of the available human capital.
Here, we test whether researchers tend to collaborate with same-gendered
colleagues, using more stringent methods and a larger dataset than in
past work. Our results reaffirm that researchers co-publish with
colleagues of the same gender more often than expected by chance, and
show that this `gender homophily' is slightly stronger today than it was
10 years ago. Contrary to our expectations, we found no evidence that
homophily is driven mostly by senior academics, and no evidence that
homophily is stronger in fields where women are in the minority.
Interestingly, journals with a high impact factor for their discipline
tended to have comparatively low homophily, as predicted if mixed-gender
teams produce better research. We discuss some potential causes of
gender homophily in science. \vspace{5mm}

\par

\noindent \textbf{Keywords:} Gender bias, Homophily, Research
collaboration, Text mining, Women in STEM.
\end{abstract}

\maketitle\unmarkedfntext{*School of BioSciences, The University of Melbourne, Victoria, Australia.}
\maketitle\unmarkedfntext{\textsuperscript{§}Organismal and Evolutionary Biology Research Programme,
Faculty of Biological and Environmental Sciences, University of Helsinki, Finland.}

@@

However, we believe that earlier studies of gender homophily were
hindered by a largely unacknowledged statistical issue that we will
refer to as the Wahlund effect (\autoref{wahlund_plot}), by analogy with
the conceptually similar Wahlund effect in population genetics {[}59{]}.
The Wahlund effect makes it deceptively difficult to infer gender-based
preferences simply by counting the number of same- and mixed-gender
coauthorships. Essentially, whenever coauthorship data are sampled from
two or more discrete sets of literature, which vary in the author gender
ratio and which are largely not connected by collaboration, the number
of same-gendered coauthors will be inflated. This can give the
impression that authors preferentially publish with same-gendered
colleagues even if no gender preferences exist, or if the true
preference is for opposite-gendered colleagues (`gender heterophily').
For example, a sample of literature containing a mixture of
bioinformatics and cell biology papers will probably contain an excess
of mostly-male and mostly-female author lists, simply because
researchers usually collaborate within their own discipline, and because
the author gender ratio is more male-biased in bioinformatics than in
cell biology {[}5{]}.

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{gender-homophily-by-discipline-time-period-and-authorship-position}{%
\subsection{Gender homophily by discipline, time period, and authorship
position}\label{gender-homophily-by-discipline-time-period-and-authorship-position}}

@@

\hypertarget{variance-in-homophily-between-disciplines}{%
\subsection{Variance in homophily between
disciplines}\label{variance-in-homophily-between-disciplines}}

@@

\hypertarget{relationship-between-gender-homophily-and-number-of-authors}{%
\subsection{Relationship between gender homophily and number of
authors}\label{relationship-between-gender-homophily-and-number-of-authors}}

Papers with two authors had significantly lower (but still positive)
\(\alpha'\) values relative to papers with more than two authors, on
average (\autoref{author_number}; statistical results in Online
Supplementary Material). Papers with 3, 4 or \(\ge5\) authors had
essentially identical average \(\alpha'\) values. The variance in
\(\alpha'\) across journals was also a little higher for 2-author papers
compared to the remainder (\autoref{author_number}), though part of this
variance is due to the reduced sample size (in terms of number of
authors) for the 2-author papers. One possible explanation for this
finding is that 2-authors papers might be more likely to have an author
list that is evenly split between career stages (e.g.~a postgraduate
student and their supervisor), increasing the chance that the authors
are mixed gender (see Figure 6). The result also suggests that the
processes responsible for gender homophily are similar in small
(e.g.~3-author) and larger (5+ author) collaborations (and across
disciplines where small versus large collaborations are the norm).

@@

\hypertarget{relationship-between-gender-homophily-and-gender-ratio}{%
\subsection{Relationship between gender homophily and gender
ratio}\label{relationship-between-gender-homophily-and-gender-ratio}}

@@

\hypertarget{relationship-between-journal-impact-factor-and-gender-homophily}{%
\subsection{Relationship between journal impact factor and gender
homophily}\label{relationship-between-journal-impact-factor-and-gender-homophily}}

@@

\hypertarget{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}{%
\subsection{Analysis accounting for differences in author gender ratio
between
countries}\label{analysis-accounting-for-differences-in-author-gender-ratio-between-countries}}

@@

\hypertarget{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}{%
\subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
the gender ratio differs between career
stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages}}

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{the-dataset}{%
\subsection{The dataset}\label{the-dataset}}

@@

\hypertarget{calculating-alpha-the-coefficient-of-homophily}{%
\subsection{\texorpdfstring{Calculating \(\alpha\), the coefficient of
homophily}{Calculating \textbackslash{}alpha, the coefficient of homophily}}\label{calculating-alpha-the-coefficient-of-homophily}}

@@

Our test assumes that the expected value of \(\alpha\) is zero if
authors randomly assort, but for small datasets this assumption is not
always true. Essentially, this issue arises because a person cannot be
their own co-author. In a small dataset comprising \(m\) men and \(f\)
women, a man can co-author with \(m - 1\) men while women can co-author
with \(m\) men. Thus, the null expectation for \(\alpha\) is a negative
number -- potentially a large one if \(m\) and \(f\) are very small.

@@

\hypertarget{minimising-the-wahlund-effect-research-discipline-and-time-period}{%
\subsection{Minimising the Wahlund effect: research discipline and time
period}\label{minimising-the-wahlund-effect-research-discipline-and-time-period}}

@@

\hypertarget{minimising-the-wahlund-effect-author-country-of-affiliation}{%
\subsection{Minimising the Wahlund effect: author country of
affiliation}\label{minimising-the-wahlund-effect-author-country-of-affiliation}}

@@

\hypertarget{calculating-standardised-journal-impact-factor}{%
\subsection{Calculating standardised journal impact
factor}\label{calculating-standardised-journal-impact-factor}}

@@

\hypertarget{statistical-analysis}{%
\subsection{Statistical analysis}\label{statistical-analysis}}

@@

\hypertarget{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}{%
\subsection{\texorpdfstring{Theoretical expectations for \(\alpha\) when
the gender ratio differs between career
stages}{Theoretical expectations for \textbackslash{}alpha when the gender ratio differs between career stages}}\label{theoretical-expectations-for-alpha-when-the-gender-ratio-differs-between-career-stages-1}}

@@

\hypertarget{data-availability-and-reproducibility}{%
\subsection{Data availability and
reproducibility}\label{data-availability-and-reproducibility}}

@@

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

@@

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Shaw_2012}{}%
1. Shaw AK, Stanton DE. Leaks in the pipeline: separating demographic
inertia from ongoing gender differences in academia. Proceedings of the
Royal Society of London B. 2012;272: 3736--3741.

\leavevmode\hypertarget{ref-Lariviere_2013}{}%
2. Larivière V, Ni C, Gingras Y, Cronin B, Sugimoto CR. Bibliometrics:
global gender disparities in science. Nature. 2013;504: 211--213.

\leavevmode\hypertarget{ref-West_2013}{}%
3. West JD, Jacquet J, King MM, Correll SJ, Bergstrom CT. The role of
gender in scholarly authorship. PLoS ONE. 2013;8: e66212.

\leavevmode\hypertarget{ref-Elsevier_report}{}%
4. Elsevier Report. Gender in the global research landscape.
elseviercom/research-intelligence/resource-library/gender-report. 2017;

\leavevmode\hypertarget{ref-Holman_2018}{}%
5. Holman L, Stuart Fox D, Hauser CE. The gender gap in science: How
long until women are equally represented? PLoS Biology. 2018;16:
e2004956.

\leavevmode\hypertarget{ref-Wutte_2007}{}%
6. Wutte M. Closing the gender gap. Nature. 2007;448: NJ101--NJ102.

\leavevmode\hypertarget{ref-Reuben_2014}{}%
7. Reuben E, Sapienza P, Zingales L. How stereotypes impair women's
careers in science. Proceedings of the National Academy of Sciences.
2014;111: 4403--4408.

\leavevmode\hypertarget{ref-Trower_2002}{}%
8. Trower CA, Chait RP. Faculty diversity: Why women and minorities are
underrepresented in the professoriate, and fresh ideas to induce needed
reform. Harvard Magazine. 2002;104: 33--37.

\leavevmode\hypertarget{ref-Umbach_2007}{}%
9. Umbach PD. Gender equity in the academic labor market: An analysis of
academic disciplines. Research in Higher Education. 2007;48: 169--192.

\leavevmode\hypertarget{ref-Hosek_2005}{}%
10. Hosek S, Cox AG, Ghosh-Dastidar B, Kofner A, Ramphal N, Scott J, et
al. Gender differences in major federal external grant programs. RAND
Corporation. 2005;

\leavevmode\hypertarget{ref-pohlhaus_2011}{}%
11. Pohlhaus JR, Jiang H, Wagner RM, Schaffer WT, Pinn VW. Sex
differences in application, success, and funding rates for NIH
extramural programs. Academic Medicine. 2011;86: 759.

\leavevmode\hypertarget{ref-Zuckerman_1987}{}%
12. Zuckerman H. Persistence and change in the careers of men and women
scientists and engineers. National Academy Press. 1987; 127--156.

\leavevmode\hypertarget{ref-Rosenfeld_1991}{}%
13. Rosenfeld RA. Outcome analysis of academic careers. Review prepared
for the Office of Scientific and Engineering Personnel, National
Research Council. 1991;

\leavevmode\hypertarget{ref-Long_1993}{}%
14. Long JS, Paul DA, Robert M. Rank advancement in academic careers:
Sex differences and the effects of productivity. American Sociological
Review. 1993; 703--722.

\leavevmode\hypertarget{ref-Hopkins_2013}{}%
15. Hopkins AL, Jawitz JW, McCarty C, Goldman A, Basu NB. Disparities in
publication patterns by gender, race and ethnicity based on a survey of
a random sample of authors. Scientometrics. 2013;96: 515--534.

\leavevmode\hypertarget{ref-ODorchai_2009}{}%
16. O'Dorchai S, Meulders D, Crippa F, Margherita A. She figures
2009--Statistics and indicators on gender equality in science.
Publications Office of the European Union. 2009;

\leavevmode\hypertarget{ref-Feldt_1986}{}%
17. Feldt B. The faculty cohort study: School of medicine. Ann Arbor,
MI: Office of Affirmative Action. 1986;

\leavevmode\hypertarget{ref-Stack_2004}{}%
18. Stack S. Gender, children and research productivity. Scientometrics.
2004;45: 891--920.

\leavevmode\hypertarget{ref-Lariviere_2011}{}%
19. Larivière V, Vignola-Gagné E, Villeneuve C, Gélinas P, Gingras Y.
Sex differences in research funding, productivity and impact: an
analysis of Québec university professors. Scientometrics. 2011;87:
483--498.

\leavevmode\hypertarget{ref-Moss_2012}{}%
20. Moss-Racusin CA, Dovidio JF, Brescoll VL, Graham MJ, Handelsman J.
Science faculty's subtle gender biases favor male students. Proceedings
of the National Academy of Sciences. 2012;109: 16474--16479.

\leavevmode\hypertarget{ref-Knobloch_2013}{}%
21. Knobloch-Westerwick S, Glynn CJ, Huge M. Science faculty's subtle
gender biases favor male students. Science Communication. 2013;35:
603--625.

\leavevmode\hypertarget{ref-Lee_2005}{}%
22. Lee S, Bozeman B. The impact of research collaboration on scientific
productivity. Social Studies of Science. 2005;35: 673--702.

\leavevmode\hypertarget{ref-Wuchty_2007}{}%
23. Wuchty S, Jones BF, Uzzi B. The increasing dominance of teams in
production of knowledge. Science. 2007;316: 1036--1039.

\leavevmode\hypertarget{ref-Abramo_2009}{}%
24. Abramo G, D'Angelo CA, Di Costa F. Research collaboration and
productivity: is there correlation? Higher Education. 2009;57: 155--171.

\leavevmode\hypertarget{ref-Lariviere_2015}{}%
25. Larivière V, Gingras Y, Sugimoto CR, Tsou A. Team size matters:
Collaboration and scientific impact since 1900. Journal of the
Association for Information Science and Technology. 2015;66: 1323--1332.

\leavevmode\hypertarget{ref-Long_1992}{}%
26. Long JS. Measures of sex differences in scientific productivity.
Social Forces. 1992;71: 159--178.

\leavevmode\hypertarget{ref-Bozeman_2011}{}%
27. Bozeman B, Monica G. How do men and women differ in research
collaborations? An analysis of the collaborative motives and strategies
of academic researchers. Research Policy. 2011;40: 1393--1402.

\leavevmode\hypertarget{ref-Abramo_2013}{}%
28. Abramo G, D'Angelo CA, Di Costa F. Gender differences in research
collaboration. Journal of Informetrics. 2013;7: 811--822.

\leavevmode\hypertarget{ref-Badar_2013}{}%
29. Badar K, Hite JM, Badir YF. Examining the relationship of
co-authorship network centrality and gender on academic research
performance: The case of chemistry researchers in pakistan.
Scientometrics. 2013;94: 755--775.

\leavevmode\hypertarget{ref-Lewison_2001}{}%
30. Lewison G. The quantity and quality of female researchers: A
bibliometric study of Iceland. Scientometrics. 2001;52: 29--43.

\leavevmode\hypertarget{ref-Webster_2001}{}%
31. Webster BM. Polish women in science: A bibliometric analysis of
Polish science and its publications. Research Evaluation. 2001;10:
185--194.

\leavevmode\hypertarget{ref-Bozeman_2004}{}%
32. Bozeman B, Corley E. Scientists' collaboration strategies:
implications for scientific and technical human capital. Research
Policy. 2004;33: 599--616.

\leavevmode\hypertarget{ref-Long_1990}{}%
33. Long JS. The origins of sex differences in science. Social Forces.
1990;68: 1297--1316.

\leavevmode\hypertarget{ref-Fuchs_2001}{}%
34. Fuchs S, Von Stebut J, Allmendinger J. Gender, science, and
scientific organizations in Germany. Minerva. 2001;39: 175--201.

\leavevmode\hypertarget{ref-Reskin_1978}{}%
35. Reskin BF. Scientific productivity, sex, and location in the
institution of science. American Journal of Sociology. 1978;83:
1235--1243.

\leavevmode\hypertarget{ref-Wright_2003}{}%
36. Wright AL, Schwindt LA, Bassford TL, Reyna VF, Shisslak PAS
Catherine M amd Germain, Reed KL. Gender differences in academic
advancement: Patterns, causes, and potential solutions in one U.S.
college of medicine. Social Forces. 2003;68: 1297--1316.

\leavevmode\hypertarget{ref-bleidorn2016age}{}%
37. Bleidorn W, Arslan RC, Denissen JJ, Rentfrow PJ, Gebauer JE, Potter
J, et al. Age and gender differences in self-esteem -- a cross-cultural
window. Journal of Personality and Social Psychology. 2016;111: 396.

\leavevmode\hypertarget{ref-jagsi_2016}{}%
38. Jagsi R, Griffith KA, Jones R, Perumalswami CR, Ubel P, Stewart A.
Sexual harassment and discrimination experiences of academic medical
faculty. JAMA. 2016;315: 2120--2121.

\leavevmode\hypertarget{ref-Martin_2014}{}%
39. Martin JL. Ten simple rules to achieve conference speaker gender
balance. PLoS computational biology. 2014;10: e1003903.

\leavevmode\hypertarget{ref-Tower_2007}{}%
40. Tower G, Julie P, Brenda R. A multidisciplinary study of
gender-based research productivity in the world's best journals. Journal
of Diversity Management. 2007;2: 23--32.

\leavevmode\hypertarget{ref-Jordan_2008}{}%
41. Jordan CE, Clark SJ, Vann CE. Do gender differences exist in the
publication productivity of accounting faculty?. Journal of Applied
Business Research. 2008;24: 77--85.

\leavevmode\hypertarget{ref-Britton_2000}{}%
42. Britton DM. The epistemology of the gendered organization. Gender
and Society. 2000;14: 418--434.

\leavevmode\hypertarget{ref-Reagans_2001}{}%
43. Reagans R, Zuckerman EW. Networks, diversity, and productivity: The
social capital of corporate R\&D teams. Organization Science. 2001;12:
502--517.

\leavevmode\hypertarget{ref-Hong_2004}{}%
44. Hong L, Page SE. Groups of diverse problem solvers can outperform
groups of high-ability problem solvers. Proceedings of the National
Academy of Sciences. 2004;101: 16385--16389.

\leavevmode\hypertarget{ref-Whittington_2008}{}%
45. Whittington KB, Smith-Doerr L. Women inventors in context:
Disparities in patenting across academia and industry. Gender \&
Society. 2008;22: 194--218.

\leavevmode\hypertarget{ref-Bear_2011}{}%
46. Bear JB, Woolley AW. The role of gender in team collaboration and
performance. Interdisciplinary Science Reviews. 2011;36: 46--153.

\leavevmode\hypertarget{ref-Herrera_2012}{}%
47. Herrera R, Duncan PA, Green MT, Skaggs SL. The effect of gender on
leadership and culture. Global Business and Organizational Excellence.
2012;31: 37--48.

\leavevmode\hypertarget{ref-Campbell_2013}{}%
48. Campbell LG, Mehtani S, Dozier ME, Rinehart J. Gender-heterogeneous
working groups produce higher quality science. PloS ONE. 2013; e79147.

\leavevmode\hypertarget{ref-Ferber_1980}{}%
49. Ferber MA, Teiman M. Are women economists at a disadvantage in
publishing journal articles? Eastern Economic Journal. 1980;6:
1189--193.

\leavevmode\hypertarget{ref-McDowell_1992}{}%
50. McDowell JM, Smith JK. The effect of gender-sorting on propensity to
coauthor: Implications for academic promotion. Economic Inquiry.
1992;30: 68--82.
